---
title: "IDS 702 Team Project"
author: "Yellow Team - Suzy Anil, Sukhpreet Sahota, Xianchi Zhang, Yuanjing Zhu"
date: '2022-10-17'
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
subtitle: 'Part II: Model and Variable Selection'
---

```{r setup, include=FALSE} 

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r, echo = FALSE, message = FALSE, results = 'hide'}
## import libraries
#install.packages('readxl')
#install.packages('dplyr')
#install.packages('tidyverse')
#install.packages('knitr')
#install.packages('ggplot2')
#install.packages('cowplot')
#install.packages('corrplot')
#install.packages('stargazer')
#install.packages('car')
#install.packages('GLMMadaptive')
#install.packages('ordinal') 
#install.packages('brant') 
#install.packages('Jmisc') 
#install.packages('caret') 

library(stargazer)
library(corrplot)
library(cowplot)
library(readxl)
library(dplyr)
library(tidyverse)
library(knitr)
library(ggplot2)
library(gridExtra)
library(MASS)
library(kableExtra)
library(tidyr)
library(broom)
library(reshape2)
library(stringr)
library(foreign)
library(nnet)
library(car)
library(GLMMadaptive)
library(ordinal)
library(brant)
library(Jmisc)
library(caret)
```


```{r, echo = FALSE, message = FALSE, results = 'hide', fig.align='center'}
# 1. Clean national income, gdp, and CO2 data - Drop any columns/country with missing values > remove them from all indicators
# Load data
inc_gdp_co2 <- read_excel('WID_Data_20102022-225315_National Income, GDP, CO2.xlsx', sheet = "Data")
dim(inc_gdp_co2)

## Delete columns with missing values
inc_gdp_co2_2 <- inc_gdp_co2 %>%
  select_if(~ !any(is.na(.)))
dim(inc_gdp_co2_2)
# here we eliminate 22 countries

year <- inc_gdp_co2_2[, "Year"]
United_States <- inc_gdp_co2_2[, grep(pattern=".USA", names(inc_gdp_co2_2), value = TRUE)]
China <- inc_gdp_co2_2[, grep(pattern=".China", names(inc_gdp_co2_2), value = TRUE)]
India <- inc_gdp_co2_2[, grep(pattern=".India", names(inc_gdp_co2_2), value = TRUE)]
Germany <- inc_gdp_co2_2[, grep(pattern=".Germany", names(inc_gdp_co2_2), value = TRUE)]
United_Kingdom <- inc_gdp_co2_2[, grep(pattern=".United Kingdom", names(inc_gdp_co2_2), value = TRUE)]
Canada <- inc_gdp_co2_2[, grep(pattern=".Canada", names(inc_gdp_co2_2), value = TRUE)]
Australia <- inc_gdp_co2_2[, grep(pattern=".Australia", names(inc_gdp_co2_2), value = TRUE)][, c(3, 6, 9)]
Brazil <- inc_gdp_co2_2[, grep(pattern=".Brazil", names(inc_gdp_co2_2), value = TRUE)]
Nigeria <- inc_gdp_co2_2[, grep(pattern=".Nigeria", names(inc_gdp_co2_2), value = TRUE)]
South_Africa <- inc_gdp_co2_2[, grep(pattern=".South Africa", names(inc_gdp_co2_2), value = TRUE)][, c(2, 4, 6)]

# combine these columns
ten_countries <- cbind(year, United_States, China, India, Germany, United_Kingdom, Canada, Australia, Brazil, Nigeria, South_Africa)
# rename the column (make it shorter...)
# rename the column (make it shorter...)
colnames(ten_countries) <- c("year", "National_income_USA" , "GDP_USA", "CO2_USA"
                             , "National_income_China" , "GDP_China", "CO2_China"
                             , "National_income_India" , "GDP_India", "CO2_India"
                             , "National_income_Germany" , "GDP_Germany", "CO2_Germany"
                             , "National_income_United_Kingdom" , "GDP_United_Kingdom", "CO2_United_Kingdom"
                             , "National_income_Canada" , "GDP_Canada", "CO2_Canada"
                             , "National_income_Australia" , "GDP_Australia", "CO2_Australia"
                             , "National_income_Brazil" , "GDP_Brazil", "CO2_Brazil"
                             , "National_income_Nigeria" , "GDP_Nigeria", "CO2_Nigeria"
                             , "National_income_South_Africa" , "GDP_South_Africa", "CO2_South_Africa")
```


```{r, echo=FALSE}
# 2a. Clean income bracket data for top 10 percent - remove data for all countries not within the selected set of 10 countries
income_bracket_top_10 <- read_excel('WID_Data_20102022-234201_Income Bracket-Top 10 Percent.xlsx', sheet = "Data")
income_bracket_top_10_percentile <- income_bracket_top_10[, "Percentile"]
income_bracket_top_10_year <- income_bracket_top_10[, "Year"]
United_States_income_bracket_10 <- income_bracket_top_10[, grep(pattern=".USA", names(income_bracket_top_10), value = TRUE)]
China_income_bracket_10 <- income_bracket_top_10[, grep(pattern=".China", names(income_bracket_top_10), value = TRUE)][, c(1)]
India_income_bracket_10 <- income_bracket_top_10[, grep(pattern=".India", names(income_bracket_top_10), value = TRUE)][, c(2)]
Germany_income_bracket_10 <- income_bracket_top_10[, grep(pattern=".Germany", names(income_bracket_top_10), value = TRUE)]
United_Kingdom_income_bracket_10 <- income_bracket_top_10[, grep(pattern=".United Kingdom", names(income_bracket_top_10), value = TRUE)]
Canada_income_bracket_10 <- income_bracket_top_10[, grep(pattern=".Canada", names(income_bracket_top_10), value = TRUE)]
Australia_income_bracket_10 <- income_bracket_top_10[, grep(pattern=".Australia", names(income_bracket_top_10), value = TRUE)]
Brazil_income_bracket_10 <- income_bracket_top_10[, grep(pattern=".Brazil", names(income_bracket_top_10), value = TRUE)]
Nigeria_income_bracket_10 <- income_bracket_top_10[, grep(pattern=".Nigeria", names(income_bracket_top_10), value = TRUE)]
South_Africa_income_bracket_10 <- income_bracket_top_10[, grep(pattern=".South Africa", names(income_bracket_top_10), value = TRUE)][, c(2)]

## Delete columns with missing values
income_bracket_top_10_ten_countries <- cbind(income_bracket_top_10_percentile, income_bracket_top_10_year, United_States_income_bracket_10, China_income_bracket_10, India_income_bracket_10, Germany_income_bracket_10, United_Kingdom_income_bracket_10, Canada_income_bracket_10, Australia_income_bracket_10, Brazil_income_bracket_10, Nigeria_income_bracket_10, South_Africa_income_bracket_10)


###################
# 2b. Clean income bracket data for mid 40 percent - remove data for all countries not within the selected set of 10 countries
income_bracket_mid_40 <- read_excel('WID_Data_20102022-234632_Income Bracket-Mid 40 Percent.xlsx', sheet = "Data")
income_bracket_mid_40_percentile <- income_bracket_mid_40[, "Percentile"]
income_bracket_mid_40_year <- income_bracket_mid_40[, "Year"]
United_States_income_bracket_40 <- income_bracket_mid_40[, grep(pattern=".USA", names(income_bracket_mid_40), value = TRUE)]
China_income_bracket_40 <- income_bracket_mid_40[, grep(pattern=".China", names(income_bracket_mid_40), value = TRUE)][, c(1)]
India_income_bracket_40 <- income_bracket_mid_40[, grep(pattern=".India", names(income_bracket_mid_40), value = TRUE)]
Germany_income_bracket_40 <- income_bracket_mid_40[, grep(pattern=".Germany", names(income_bracket_mid_40), value = TRUE)]
United_Kingdom_income_bracket_40 <- income_bracket_mid_40[, grep(pattern=".United Kingdom", names(income_bracket_mid_40), value = TRUE)]
Canada_income_bracket_40 <- income_bracket_mid_40[, grep(pattern=".Canada", names(income_bracket_mid_40), value = TRUE)]
Australia_income_bracket_40 <- income_bracket_mid_40[, grep(pattern=".Australia", names(income_bracket_mid_40), value = TRUE)]
Brazil_income_bracket_40 <- income_bracket_mid_40[, grep(pattern=".Brazil", names(income_bracket_mid_40), value = TRUE)]
Nigeria_income_bracket_40 <- income_bracket_mid_40[, grep(pattern=".Nigeria", names(income_bracket_mid_40), value = TRUE)]
South_Africa_income_bracket_40 <- income_bracket_mid_40[, grep(pattern=".South Africa", names(income_bracket_mid_40), value = TRUE)][, c(2)]

## Delete columns with missing values
income_bracket_mid_40_ten_countries <- cbind(income_bracket_mid_40_percentile, income_bracket_mid_40_year, United_States_income_bracket_40, China_income_bracket_40, India_income_bracket_40, Germany_income_bracket_40, United_Kingdom_income_bracket_40, Canada_income_bracket_40, Australia_income_bracket_40, Brazil_income_bracket_40, Nigeria_income_bracket_40, South_Africa_income_bracket_40)


###################
# 2c. Clean income bracket data for bottom 50 percent - remove data for all countries not within the selected set of 10 countries
income_bracket_bottom_50 <- read_excel('WID_Data_20102022-234329_Income Bracket-Bottom 50 Percent.xlsx', sheet = "Data")
income_bracket_bottom_50_percentile <- income_bracket_bottom_50[, "Percentile"]
income_bracket_bottom_50_year <- income_bracket_bottom_50[, "Year"]
United_States_income_bracket_50 <- income_bracket_bottom_50[, grep(pattern=".USA", names(income_bracket_bottom_50), value = TRUE)]
China_income_bracket_50 <- income_bracket_bottom_50[, grep(pattern=".China", names(income_bracket_bottom_50), value = TRUE)][, c(1)]
India_income_bracket_50 <- income_bracket_bottom_50[, grep(pattern=".India", names(income_bracket_bottom_50), value = TRUE)]
Germany_income_bracket_50 <- income_bracket_bottom_50[, grep(pattern=".Germany", names(income_bracket_bottom_50), value = TRUE)]
United_Kingdom_income_bracket_50 <- income_bracket_bottom_50[, grep(pattern=".United Kingdom", names(income_bracket_bottom_50), value = TRUE)]
Canada_income_bracket_50 <- income_bracket_bottom_50[, grep(pattern=".Canada", names(income_bracket_bottom_50), value = TRUE)]
Australia_income_bracket_50 <- income_bracket_bottom_50[, grep(pattern=".Australia", names(income_bracket_bottom_50), value = TRUE)]
Brazil_income_bracket_50 <- income_bracket_bottom_50[, grep(pattern=".Brazil", names(income_bracket_bottom_50), value = TRUE)]
Nigeria_income_bracket_50 <- income_bracket_bottom_50[, grep(pattern=".Nigeria", names(income_bracket_bottom_50), value = TRUE)]
South_Africa_income_bracket_50 <- income_bracket_bottom_50[, grep(pattern=".South Africa", names(income_bracket_bottom_50), value = TRUE)][, c(2)]

## Delete columns with missing values
income_bracket_bottom_50_ten_countries <- cbind(income_bracket_bottom_50_percentile, income_bracket_bottom_50_year, United_States_income_bracket_50, China_income_bracket_50, India_income_bracket_50, Germany_income_bracket_50, United_Kingdom_income_bracket_50, Canada_income_bracket_50, Australia_income_bracket_50, Brazil_income_bracket_50, Nigeria_income_bracket_50, South_Africa_income_bracket_50)


###################
# 2d. Combine Top 10, Mid 40, Bottom 50 tables
income_bracket_all_ten_countries <- cbind(income_bracket_top_10_year, United_States_income_bracket_10, United_States_income_bracket_40, United_States_income_bracket_50, China_income_bracket_10, China_income_bracket_40, China_income_bracket_50, India_income_bracket_10, India_income_bracket_40, India_income_bracket_50, Germany_income_bracket_10, Germany_income_bracket_40, Germany_income_bracket_50, United_Kingdom_income_bracket_10, United_Kingdom_income_bracket_40, United_Kingdom_income_bracket_50, Canada_income_bracket_10, Canada_income_bracket_40, Canada_income_bracket_50, Australia_income_bracket_10, Australia_income_bracket_40, Australia_income_bracket_50, Brazil_income_bracket_10, Brazil_income_bracket_40, Brazil_income_bracket_50, Nigeria_income_bracket_10, Nigeria_income_bracket_40, Nigeria_income_bracket_50, South_Africa_income_bracket_10, South_Africa_income_bracket_40, South_Africa_income_bracket_50)
```


```{r, echo=FALSE}
# 3. Clean population data - remove data for all countries not within the selected set of 10 countries
population_data <- read_excel('WID_Data_20102022-234748_Population.xlsx', sheet = "Data")
population_percentile <- population_data[, "Percentile"]
population_year <- population_data[, "Year"]
United_States_population <- population_data[, grep(pattern=".USA", names(population_data), value = TRUE)]
China_population <- population_data[, grep(pattern=".China", names(population_data), value = TRUE)][, c(1)]
India_population <- population_data[, grep(pattern=".India", names(population_data), value = TRUE)]
Germany_population <- population_data[, grep(pattern=".Germany", names(population_data), value = TRUE)]
United_Kingdom_population <- population_data[, grep(pattern=".United Kingdom", names(population_data), value = TRUE)]
Canada_population <- population_data[, grep(pattern=".Canada", names(population_data), value = TRUE)]
Australia_population <- population_data[, grep(pattern=".Australia", names(population_data), value = TRUE)]
Brazil_population <- population_data[, grep(pattern=".Brazil", names(population_data), value = TRUE)]
Nigeria_population <- population_data[, grep(pattern=".Nigeria", names(population_data), value = TRUE)]
South_Africa_population <- population_data[, grep(pattern=".South Africa", names(population_data), value = TRUE)][, c(2)]

## Delete columns not within selected set of 10 countries
population_ten_countries <- cbind(population_percentile, population_year, United_States_population, China_population, India_population, Germany_population, United_Kingdom_population, Canada_population, Australia_population, Brazil_population, Nigeria_population, South_Africa_population)
```


```{r, echo=FALSE}
# 4. Clean average wealth data - remove data for all countries not within the selected set of 10 countries
wealth_data <- read_excel('WID_Data_20102022-234942_Average Wealth-Per Adult, Wealth Income Ratio.xlsx', sheet = "Data")
wealth_percentile <- wealth_data[, "Percentile"]
wealth_year <- wealth_data[, "Year"]
United_States_wealth <- wealth_data[, grep(pattern=".USA", names(wealth_data), value = TRUE)]
China_wealth <- wealth_data[, grep(pattern=".China", names(wealth_data), value = TRUE)]
India_wealth <- wealth_data[, grep(pattern=".India", names(wealth_data), value = TRUE)]
Germany_wealth <- wealth_data[, grep(pattern=".Germany", names(wealth_data), value = TRUE)]
United_Kingdom_wealth <- wealth_data[, grep(pattern=".United Kingdom", names(wealth_data), value = TRUE)]
Canada_wealth <- wealth_data[, grep(pattern=".Canada", names(wealth_data), value = TRUE)]
Australia_wealth <- wealth_data[, grep(pattern=".Australia", names(wealth_data), value = TRUE)][, c(3, 8)]
Brazil_wealth <- wealth_data[, grep(pattern=".Brazil", names(wealth_data), value = TRUE)]
Nigeria_wealth <- wealth_data[, grep(pattern=".Nigeria", names(wealth_data), value = TRUE)]
South_Africa_wealth <- wealth_data[, grep(pattern=".South Africa", names(wealth_data), value = TRUE)][, c(2, 5)]

## Delete columns not within selected set of 10 countries
wealth_ten_countries <- cbind(wealth_percentile, wealth_year, United_States_wealth, China_wealth, India_wealth, Germany_wealth, United_Kingdom_wealth, Canada_wealth, Australia_wealth, Brazil_wealth, Nigeria_wealth, South_Africa_wealth)
```


```{r, echo=FALSE, result = "hide"}
#correlation matrix
only_co2_ten_countries <- ten_countries[, c(4,7,10,13,16,19,22,25,28,31)]
only_wealth_ten_countries <- wealth_ten_countries[, c(3,5,7,9,11,13,15,17,19,21)]
co2_wealth_ten_countries <- cbind(only_co2_ten_countries, only_wealth_ten_countries)
colnames(co2_wealth_ten_countries) <- c(
  "co2_usa","co2_china","co2_india","co2_germany","co2_uk","co2_canada","co2_australia","co2_brazil","co2_nigeria","co2_south africa", 
  "wealth_usa","wealth_china","wealth_india","wealth_germany","wealth_uk","wealth_canada","wealth_australia","wealth_brazil","wealth_nigeria","wealth_south africa")

#corrplot(cor(co2_wealth_ten_countries), method  = "circle", type = "upper")
#cor(co2_wealth_ten_countries)
```


### Model 1
```{r, echo=FALSE,message = FALSE, result = "hide"}

# stacking income brackets
income_bracket_USA <- income_bracket_all_ten_countries[, c(2,3,4)]
colnames(income_bracket_USA) <- c('top10','middle40','bottom50')
income_bracket_china <- income_bracket_all_ten_countries[, c(5,6,7)]
colnames(income_bracket_china) <- c('top10','middle40','bottom50')
income_bracket_India <- income_bracket_all_ten_countries[, c(8,9,10)]
colnames(income_bracket_India) <- c('top10','middle40','bottom50')
income_bracket_Germany <- income_bracket_all_ten_countries[, c(11,12,13)]
colnames(income_bracket_Germany) <- c('top10','middle40','bottom50')
income_bracket_UK <- income_bracket_all_ten_countries[, c(14,15,16)]
colnames(income_bracket_UK) <- c('top10','middle40','bottom50')
income_bracket_Canada <- income_bracket_all_ten_countries[, c(17,18,19)]
colnames(income_bracket_Canada) <- c('top10','middle40','bottom50')
income_bracket_Australia <- income_bracket_all_ten_countries[, c(20,21,22)]
colnames(income_bracket_Australia) <- c('top10','middle40','bottom50')
income_bracket_Brazil <- income_bracket_all_ten_countries[, c(23,24,25)]
colnames(income_bracket_Brazil) <- c('top10','middle40','bottom50')
income_bracket_Nigeria <- income_bracket_all_ten_countries[, c(26,27,28)]
colnames(income_bracket_Nigeria) <- c('top10','middle40','bottom50')
income_bracket_NorthAfrica <- income_bracket_all_ten_countries[, c(29,30,31)]
colnames(income_bracket_NorthAfrica) <- c('top10','middle40','bottom50')

income_brackets <- bind_rows(income_bracket_USA, income_bracket_china, income_bracket_India, income_bracket_Germany, income_bracket_UK, income_bracket_Canada, income_bracket_Australia, income_bracket_Brazil, income_bracket_Nigeria, income_bracket_NorthAfrica)
country_code <- rep(1:10, each = 21)
income_brackets_for_model <- cbind(income_brackets, country_code)
income_brackets_for_model$country_code <- factor(income_brackets_for_model$country_code)

# stacking carbon footprint
co2_USA <- co2_wealth_ten_countries[, 1]
co2_China <- co2_wealth_ten_countries[, 2]
co2_India <- co2_wealth_ten_countries[, 3]
co2_Germany <- co2_wealth_ten_countries[, 4]
co2_UK <- co2_wealth_ten_countries[, 5]
co2_Canada <- co2_wealth_ten_countries[, 6]
co2_Australia <- co2_wealth_ten_countries[, 7]
co2_Brazil <- co2_wealth_ten_countries[, 8]
co2_Nigira <- co2_wealth_ten_countries[, 9]
co2_SouthAfrica <- co2_wealth_ten_countries[, 10]

stacking_co2 <- c(co2_USA,co2_China,co2_India,co2_Germany,co2_UK, co2_Canada, co2_Australia, co2_Brazil, co2_Nigira, co2_SouthAfrica)

dataset_for_model1 <- cbind(income_brackets_for_model, stacking_co2)

```


```{r, echo=FALSE,message = FALSE, result = "hide"}
mlr_q1 <- lm(stacking_co2 ~ top10 + middle40 + bottom50 + country_code, data = dataset_for_model1)
summary(mlr_q1)

mlr_q2 <- lm(stacking_co2 ~ top10 + middle40 + bottom50, data = dataset_for_model1)
summary(mlr_q2)

mlr_q3 <- lm(log(stacking_co2) ~ top10 + middle40 + bottom50, data = dataset_for_model1)
summary(mlr_q3)

# backward AIC
mod <- lm(stacking_co2 ~.,data = dataset_for_model1)
mod.back <- stepAIC(mod,direction="backward")
summary(mod.back)

```

```{r}
summary(mlr_q1)
plot(mlr_q1)
```

```{r}
summary(mlr_q2)
plot(mlr_q2)
```

```{r}
summary(mlr_q3)
plot(mlr_q3)
```
The baseline of country is USA. Top 10% income bracket and country are statistically significant. 


```{r,echo = FALSE, message = FALSE, fig.align='center'}

mlr_q1_final <- lm(stacking_co2 ~ top10 + country_code,data = dataset_for_model1)
summary(mlr_q1_final)

mlr_q1_final %>%
  tidy() %>% 
  kable(caption = "MLR Model Regressing Carbon Footprint on Income Inequality",
        col.names = c("Predictor", "Estimate", "SE", "t", "p-value"),
        digits = c(2, 2, 2, 2, 2),
        align = "l") %>% 
  add_footnote(c("Multiple R-squared: 0.9160","Adjusted R-squared: 0.9118"))%>% 
  kable_styling(position="center", full_width = T)
```
```{r,echo = FALSE, message = FALSE, fig.align='center'}
confint(mlr_q1_final, level = 0.95) %>%  
  kable(caption="95% Confidence Interval",
         digits = c(2, 2),
         align = "l") %>%
  kable_styling(position="center", full_width = T)
```


```{r}
summary(mlr_q1_final)
plot(mlr_q1_final)
```

```{r echo = FALSE, message = FALSE, fig.width = 12, fig.height = 12, fig.align = 'center'}
# check assumptions
par(mfrow = c(3, 2))
plot(mlr_q1_final, which = 1)
plot(mlr_q1_final, which = 2)
plot(mlr_q1_final, which = 3)
plot(mlr_q1_final, which = 4)
plot(mlr_q1_final, which = 5)

```

```{r}

dataset_for_model1$top10demean <- demean(dataset_for_model1$top10)
dataset_for_model1$midde40demean <- demean(dataset_for_model1$middle40)
dataset_for_model1$bottom50demean <- demean(dataset_for_model1$bottom50)

mlr_q4 <- lm(stacking_co2 ~ top10demean + midde40demean + bottom50demean, data = dataset_for_model1)
summary(mlr_q4)
```

```{r}
plot(mlr_q4)
```


```{r}
mlr_q5 <- lm(log(stacking_co2) ~ top10demean + midde40demean + bottom50demean, data = dataset_for_model1)
summary(mlr_q5)
plot(mlr_q5)
```

```{r}
dataset_for_model1$top10log <- log(dataset_for_model1$top10)
dataset_for_model1$midde40log <- log(dataset_for_model1$middle40)
dataset_for_model1$bottom50log <- log(dataset_for_model1$bottom50)

mlr_q6 <- lm(stacking_co2 ~ top10log + midde40log + bottom50log, data = dataset_for_model1)
summary(mlr_q6)
plot(mlr_q6)
```

```{r}
mlr_q7 <- lm(log(stacking_co2) ~ top10log + midde40log + bottom50log, data = dataset_for_model1)
summary(mlr_q7)
plot(mlr_q7)
```


## Model 2

```{r}
ten_countries
population_ten_countries
co2_wealth_ten_countries
# meanC02 <- mean(as.matrix(only_co2_ten_countries))
# meanC02
```
```{r}
colnames(population_ten_countries) <- c("Percentile","Year","pop_usa","pop_china","pop_india","pop_germany","pop_uk","pop_canada","pop_aus_nz","pop_oceania", "pop_australia","pop_brazil","pop_nigeria","pop_southafrica")
population_ten_countries <- subset(population_ten_countries, select=-c(pop_aus_nz, pop_oceania))
population_ten_countries
```

```{r}
nat_inc <- ten_countries[ , grepl("National_income", names( ten_countries ) ) ]
nat_inc$Year <- 2000:2020
co2 <- ten_countries[ , grepl("CO2", names( ten_countries ) ) ]
co2$Year <- 2000:2020
pop <- population_ten_countries[ , grepl("pop", names( population_ten_countries ) ) ]
pop$Year <- 2000:2020
wealth <- co2_wealth_ten_countries[ , grepl("wealth", names( co2_wealth_ten_countries ) ) ]
wealth$Year <- 2000:2020
co2
```
```{r}
wealth_stack <- reshape2::melt(wealth, id.vars = "Year", variable.name = 'country_wealth', value.name='National_Wealth')
wealth_stack$Country <- toupper(sub(".*_", "", wealth_stack$country_wealth))
pop_stack <- reshape2::melt(pop, id.vars = "Year", variable.name='country_pop', value.name='National_Population')
pop_stack$Country <- toupper(sub(".*_", "", pop_stack$country_pop))
pop_stack$Country[pop_stack$Country == "SOUTHAFRICA"] <- "SOUTH AFRICA"
co2_stack <- reshape2::melt(co2, id.vars = "Year", variable.name='country_co2', value.name='National_Carbon_Emissions')
co2_stack$Country <- toupper(sub(".*_", "", co2_stack$country_co2))
co2_stack$Country[co2_stack$Country == "KINGDOM"] <- "UK"
co2_stack$Country[co2_stack$Country == "AFRICA"] <- "SOUTH AFRICA"
nat_inc_stack <- reshape2::melt(nat_inc, id.vars = "Year", variable.name='country_income', value.name='National_Income')
nat_inc_stack$Country <- toupper(sub(".*_", "", nat_inc_stack$country_income))
nat_inc_stack$Country[nat_inc_stack$Country == "KINGDOM"] <- "UK"
nat_inc_stack$Country[nat_inc_stack$Country == "AFRICA"] <- "SOUTH AFRICA"
complete_model_2_df <- wealth_stack[, c("Year", "Country", "National_Wealth")]
complete_model_2_df <- merge(complete_model_2_df, pop_stack, by=c("Year", "Country"))
complete_model_2_df <- merge(complete_model_2_df, co2_stack, by=c("Year", "Country"))
complete_model_2_df <- merge(complete_model_2_df, nat_inc_stack, by=c("Year", "Country"))
dropcolFromCompleteModel <- c("country_pop", "country_co2", "country_income")
complete_model_2_df = complete_model_2_df[, !(names(complete_model_2_df) %in% dropcolFromCompleteModel)]
complete_model_2_df <- complete_model_2_df[order(complete_model_2_df[,2], complete_model_2_df[,1]), ]
```

```{r}
mean_co2 <- mean(co2_stack$National.carbon.emissions)
```


```{r, echo=FALSE}
##### create the dataset
Year <- population_ten_countries["Year"]
Country <- co2_stack["Country"]
CO2<- co2_stack["National_Carbon_Emissions"]
National_wealth <- wealth_stack["National_Wealth"]
National_income<- nat_inc_stack["National_Income"]
Population <- pop_stack["National_Population"]
m2_data<- data.frame(Year, Country, CO2, National_wealth, National_income, Population)
colnames(m2_data)
names(m2_data)[names(m2_data) == "National_Carbon_Emissions"] <- "CO2"
names(m2_data)[names(m2_data) == "National_Population"] <- "Population"
# m2_data$Country <- factor(m2_data$Country)
meanC02 <- mean(as.matrix(only_co2_ten_countries))
meanC02
m2_data1<-m2_data %>% mutate(
  CO2_C = case_when(CO2 < 500 ~ "low",
                    CO2<2000 ~ "med",
                    CO2 >= 2000 ~ "high")
)
m2_data1$Country <- factor(m2_data1$Country)
m2_data1$CO2_C <- factor(m2_data1$CO2_C)
m2_data1$CO2_C <- ordered(m2_data1$CO2_C, levels = c("low", "med", "high"))
str(m2_data1)
m2_data1
```
```{r, echo=FALSE, message = FALSE, fig.width = 12, fig.height = 10, fig.align = 'center'}
##### EDA analysis for model 2
# colorblind-friendly palette
colorBlindBlack3  <- c("#009E73", "#F0E442", "#D55E00")

## CO2-income
eda_m2_1 <- ggplot(m2_data1, aes(x=CO2_C, y = National_Income, fill = CO2_C)) +
  geom_boxplot() +
  stat_summary(fun.y="mean") +
  scale_fill_manual(values=colorBlindBlack3) +
  labs(title = "CO2 vs Income",
       x = "Carbon Emission Level",
       y = "National Income",
       fill = "CO2 Level") +
  theme_classic() + theme(legend.position = "top")

## CO2-wealth
eda_m2_2 <- ggplot(m2_data1, aes(x=CO2_C, y = National_Wealth, fill = CO2_C)) +
  geom_boxplot() +
  stat_summary(fun.y="mean") +
  scale_fill_manual(values=colorBlindBlack3) +
  labs(title = "CO2 vs Wealth",
       x = "Carbon Emission Level",
       y = "National Wealth",
       fill = "CO2 Level") +
  theme_classic() + theme(legend.position = "top")

## CO2-Population
eda_m2_3 <- ggplot(m2_data1, aes(x=CO2_C, y = Population, fill = CO2_C)) +
  geom_boxplot() +
  stat_summary(fun.y="mean") +
  scale_fill_manual(values=colorBlindBlack3) +
  labs(title = "CO2 vs Population",
       x = "Carbon Emission Level",
       y = "National population",
       fill = "CO2 Level") +
  theme_classic() + theme(legend.position = "top")

grid.arrange(eda_m2_1,eda_m2_2,eda_m2_3, nrow = 2)

```

```{r, echo=FALSE}
##### create the model2
m2 <- multinom(CO2_C ~ Country + National_Wealth + National_Income + Population, data = m2_data1, Hess = TRUE)
summary(m2)
### the CO2_C = low  is the base line. 
exp(coef(m2))
confint(m2)
kable(exp(confint(m2)))
output <- summary(m2)
z_value <- output$coefficients/output$standard.errors
p_value <- (1- pnorm(abs(z_value), 0, 1))*2
full_summary <- lapply(c("med", "high"), function(x) rbind(output$coefficients[as.character(x),],
                                                                    output$standard.errors[as.character(x),],
                                                                    z_value[as.character(x),],
                                                                    p_value[as.character(x),]))

kable(lapply(full_summary,function(x) {rownames(x) <- c("Coefficient", "Std.Errors", "z-value", "p-value");x}))
```
This is a TEST

```{r, echo=FALSE}
summary(m2)
```


From the p-value table below, it shows that every variable in the model are significant. So let's do a deviance test. 
```{r, echo=FALSE}
##### Deviance test
##### test Country
m2noc <- multinom(CO2_C ~National_Wealth + National_Income + Population, data = m2_data1)
anova(m2,m2noc, test = "Chisq")
# p-value is less than 0.05, so country is good. 
##### test national wealth
m2nonw <- multinom(CO2_C ~ Country + National_Income + Population, data = m2_data1)
anova(m2,m2nonw, test = "Chisq")
#### p-value is larger than 0.05, so wealth is not good. 

##### test national wealth * country
m2nonw1 <- multinom(CO2_C ~ Country*National_Wealth + National_Income+Population, data = m2_data1)
anova(m2,m2nonw1, test = "Chisq")
#### p-value is larger than 0.05, so national wealth*country is not good. 

##### test national_income
m2noni <- multinom(CO2_C ~ Country+National_Wealth+ Population, data = m2_data1)
anova(m2,m2noni, test = "Chisq")
#### p-value is larger than 0.05, so national income is not good. 

##### test national_income*country
m2noni1 <- multinom(CO2_C ~ National_Wealth+ Country*National_Income+ Population, data = m2_data1)
anova(m2,m2noni1, test = "Chisq")
#### p-value is larger than 0.05, so national income interaction is not good. 

##### test population
m2nop <- multinom(CO2_C ~ Country+National_Wealth+ National_Income, data = m2_data1)
anova(m2,m2nop, test = "Chisq")
#### p-value is less than 0.05, so population is good. 

##### test population interaction
m2nop1 <- multinom(CO2_C ~ Country + National_Wealth+ National_Income+Population*Country, data = m2_data1)
anova(m2,m2nop1, test = "Chisq")
#### p-value is larger than 0.05, so population interaction is not good. 
```
so it shows only  and country are significant after deviance test.
```{r, echo=FALSE}
# backward AIC
#mod <- multinom(CO2_C ~.,data = m2_data1)
#mod.back <- stepAIC(mod,direction="backward")
#summary(mod.back)

```


```{r}
##mixed_model(fixed = CO2_C ~ National_Wealth + National_Income + Population*Country, random = ~ Country | Population, data=m2_data1, family=binomial())
```

```{r}
m2_data1$pop_log <- log(m2_data1$Population)
m2_data1$income_log <- log(m2_data1$National_Income)
m2_data1$wealth_log <- log(m2_data1$National_Wealth)
m2_data1
```

controlling for country... just the name of the country does not give use information
```{r}
mixcl <- clmm(CO2_C ~ income_log + wealth_log + pop_log + (1|Country), data=m2_data1)
summary(mixcl)
```

```{r}
#vif(mixcl)
```


```{r}
multinomM2 <- polr(CO2_C ~ income_log + wealth_log + pop_log, data=m2_data1, Hess = TRUE)
summary(multinomM2)
brant(multinomM2)
```


### Predict 10 countries

```{r, echo = FALSE, message = FALSE, results = 'hide', fig.align='center'}
## select 10 countries
yearPredict <- inc_gdp_co2_2[, "Year"]
France <- inc_gdp_co2_2[, grep(pattern=".France", names(inc_gdp_co2_2), value = TRUE)]
Japan <- inc_gdp_co2_2[, grep(pattern=".Japan", names(inc_gdp_co2_2), value = TRUE)]
Russia <- inc_gdp_co2_2[, grep(pattern=".Russian Federation", names(inc_gdp_co2_2), value = TRUE)]
Mexico <- inc_gdp_co2_2[, grep(pattern=".Mexico", names(inc_gdp_co2_2), value = TRUE)]
Belgium <- inc_gdp_co2_2[, grep(pattern=".Belgium", names(inc_gdp_co2_2), value = TRUE)]
Norway <- inc_gdp_co2_2[, grep(pattern=".Norway", names(inc_gdp_co2_2), value = TRUE)]
Indonesia <- inc_gdp_co2_2[, grep(pattern=".Indonesia", names(inc_gdp_co2_2), value = TRUE)]
Argentina <- inc_gdp_co2_2[, grep(pattern=".Argentina", names(inc_gdp_co2_2), value = TRUE)]
Ghana <- inc_gdp_co2_2[, grep(pattern=".Ghana", names(inc_gdp_co2_2), value = TRUE)]
Qatar <- inc_gdp_co2_2[, grep(pattern=".Qatar", names(inc_gdp_co2_2), value = TRUE)]

# combine these columns
ten_countries_predict <- cbind(yearPredict, France, Japan, Russia, Mexico, Belgium, Norway, Indonesia, Argentina, Ghana, Qatar)
# rename the column (make it shorter...)
# rename the column (make it shorter...)
colnames(ten_countries_predict) <- c("year", "National_income_France" , "GDP_France", "CO2_France"
                             , "National_income_Japan" , "GDP_Japan", "CO2_Japan"
                             , "National_income_Russia" , "GDP_Russia", "CO2_Russia"
                             , "National_income_Mexico" , "GDP_Mexico", "CO2_Mexico"
                             , "National_income_Belgium" , "GDP_Belgium", "CO2_Belgium"
                             , "National_income_Norway" , "GDP_Norway", "CO2_Norway"
                             , "National_income_Indonesia" , "GDP_Indonesia", "CO2_Indonesia"
                             , "National_income_Argentina" , "GDP_Argentina", "CO2_Argentina"
                             , "National_income_Ghana" , "GDP_Ghana", "CO2_Ghana"
                             , "National_income_Qatar" , "GDP_Qatar", "CO2_Qatar")

```

```{r, echo=FALSE}
# 3. Clean population data - remove data for all countries not within the selected set of 10 countries

population_percentile_predict <- population_data[, "Percentile"]
population_year_predict <- population_data[, "Year"]
France_population_predict <- population_data[, grep(pattern=".France", names(population_data), value = TRUE)]
Japan_population_predict <- population_data[, grep(pattern=".Japan", names(population_data), value = TRUE)][, c(1)]
Russia_population_predict <- population_data[, grep(pattern=".Russian Federation", names(population_data), value = TRUE)]
Mexico_population_predict <- population_data[, grep(pattern=".Mexico", names(population_data), value = TRUE)]
Belgium_population_predict <- population_data[, grep(pattern=".Belgium", names(population_data), value = TRUE)]
Norway_population_predict <- population_data[, grep(pattern=".Norway", names(population_data), value = TRUE)]
Indonesia_population_predict <- population_data[, grep(pattern=".Indonesia", names(population_data), value = TRUE)]
Argentina_population_predict <- population_data[, grep(pattern=".Argentina", names(population_data), value = TRUE)]
Ghana_population_predict <- population_data[, grep(pattern=".Ghana", names(population_data), value = TRUE)]
Qatar_population_predict <- population_data[, grep(pattern=".Qatar", names(population_data), value = TRUE)]

## Delete columns not within selected set of 10 countries
population_ten_countries_predict <- cbind(population_percentile_predict, population_year_predict, France_population_predict, Japan_population_predict, Russia_population_predict, Mexico_population_predict, Belgium_population_predict, Norway_population_predict, Indonesia_population_predict, Argentina_population_predict, Ghana_population_predict, Qatar_population_predict)
```


```{r, echo=FALSE}
# 4. Clean average wealth data - remove data for all countries not within the selected set of 10 countries

wealth_percentile_predict <- wealth_data[, "Percentile"]
wealth_year_predict <- wealth_data[, "Year"]
France_wealth_predict <- wealth_data[, grep(pattern=".France", names(wealth_data), value = TRUE)]
Japan_wealth_predict <- wealth_data[, grep(pattern=".Japan", names(wealth_data), value = TRUE)]
Russia_wealth_predict <- wealth_data[, grep(pattern=".Russian Federation", names(wealth_data), value = TRUE)]
Mexico_wealth_predict <- wealth_data[, grep(pattern=".Mexico", names(wealth_data), value = TRUE)]
Belgium_wealth_predict <- wealth_data[, grep(pattern=".Belgium", names(wealth_data), value = TRUE)]
Norway_wealth_predict <- wealth_data[, grep(pattern=".Norway", names(wealth_data), value = TRUE)]
Indonesia_wealth_predict <- wealth_data[, grep(pattern=".Indonesia", names(wealth_data), value = TRUE)]
Argentina_wealth_predict <- wealth_data[, grep(pattern=".Argentina", names(wealth_data), value = TRUE)]
Ghana_wealth_predict <- wealth_data[, grep(pattern=".Ghana", names(wealth_data), value = TRUE)]
Qatar_wealth_predict <- wealth_data[, grep(pattern=".Qatar", names(wealth_data), value = TRUE)]

## Delete columns not within selected set of 10 countries
wealth_ten_countries_predict <- cbind(wealth_percentile_predict, wealth_year_predict, France_wealth_predict, Japan_wealth_predict, Russia_wealth_predict, Mexico_wealth_predict, Belgium_wealth_predict, Norway_wealth_predict, Indonesia_wealth_predict, Argentina_wealth_predict, Ghana_wealth_predict, Qatar_wealth_predict)
```

```{r, echo=FALSE, result = "hide"}
#correlation matrix
only_co2_ten_countries_predict <- ten_countries_predict[, c(4,7,10,13,16,19,22,25,28,31)]
only_wealth_ten_countries_predict <- wealth_ten_countries_predict[, c(3,5,7,9,11,13,15,17,19,21)]
co2_wealth_ten_countries_predict <- cbind(only_co2_ten_countries_predict, only_wealth_ten_countries_predict)
colnames(co2_wealth_ten_countries_predict) <- c(
  "co2_france","co2_japan","co2_russia","co2_mexico","co2_belgium","co2_norway","co2_indonesia","co2_argentina","co2_ghana","co2_qatar", 
  "wealth_france","wealth_japan","wealth_russia","wealth_mexico","wealth_belgium","wealth_norway","wealth_indonesia","wealth_argentina","wealth_ghana","wealth_qatar")

#corrplot(cor(co2_wealth_ten_countries), method  = "circle", type = "upper")
#cor(co2_wealth_ten_countries)
```

```{r}
colnames(population_ten_countries_predict) <- c("Percentile","Year","pop_france","pop_japan","pop_russia","pop_mexico","pop_belgium","pop_norway","pop_indonesia","pop_argentina", "pop_ghana","pop_qatar")
population_ten_countries_predict
```

```{r}
nat_inc_predict <- ten_countries_predict[ , grepl("National_income", names( ten_countries_predict ) ) ]
nat_inc_predict$Year <- 2000:2020
co2_predict <- ten_countries_predict[ , grepl("CO2", names( ten_countries_predict ) ) ]
co2_predict$Year <- 2000:2020
pop_predict <- population_ten_countries_predict[ , grepl("pop", names( population_ten_countries_predict ) ) ]
pop_predict$Year <- 2000:2020
wealth_predict <- co2_wealth_ten_countries_predict[ , grepl("wealth", names( co2_wealth_ten_countries_predict ) ) ]
wealth_predict$Year <- 2000:2020
```

```{r}
wealth_stack_predict <- reshape2::melt(wealth_predict, id.vars = "Year", variable.name = 'country_wealth', value.name='National_Wealth')
wealth_stack_predict$Country <- toupper(sub(".*_", "", wealth_stack_predict$country_wealth))
pop_stack_predict <- reshape2::melt(pop_predict, id.vars = "Year", variable.name='country_pop', value.name='National_Population')
pop_stack_predict$Country <- toupper(sub(".*_", "", pop_stack_predict$country_pop))
co2_stack_predict <- reshape2::melt(co2_predict, id.vars = "Year", variable.name='country_co2', value.name='National_Carbon_Emissions')
co2_stack_predict$Country <- toupper(sub(".*_", "", co2_stack_predict$country_co2))
nat_inc_stack_predict <- reshape2::melt(nat_inc_predict, id.vars = "Year", variable.name='country_income', value.name='National_Income')
nat_inc_stack_predict$Country <- toupper(sub(".*_", "", nat_inc_stack_predict$country_income))
predict_model_df <- wealth_stack_predict[, c("Year", "Country", "National_Wealth")]
predict_model_df <- merge(predict_model_df, pop_stack_predict, by=c("Year", "Country"))
predict_model_df <- merge(predict_model_df, co2_stack_predict, by=c("Year", "Country"))
predict_model_df <- merge(predict_model_df, nat_inc_stack_predict, by=c("Year", "Country"))
dropcolFromPredictModel <- c("country_pop", "country_co2", "country_income")
predict_model_df = predict_model_df[, !(names(predict_model_df) %in% dropcolFromPredictModel)]
predict_model_df <- predict_model_df[order(predict_model_df[,2], predict_model_df[,1]), ]
```

```{r}
mixcl2 <- clmm2(CO2_C ~ income_log + wealth_log + pop_log, random = Country, data=m2_data1, Hess=TRUE)
summary(mixcl2)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis', header=FALSE, quiet = TRUE, include=FALSE}
rownames(predict_model_df) <- 1:nrow(predict_model_df)
predict_model_df$pop_log <- log(predict_model_df$National_Population)
predict_model_df$income_log <- log(predict_model_df$National_Income)
predict_model_df$wealth_log <- log(predict_model_df$National_Wealth)
predict_model_df <- predict_model_df %>% mutate(
  CO2_C = case_when(National_Carbon_Emissions < 500 ~ "low",
                    National_Carbon_Emissions<2000 ~ "med",
                    National_Carbon_Emissions >= 2000 ~ "high"))
predict_model_df$CO2_C <- factor(predict_model_df$CO2_C)
predict_model_df$CO2_C <- ordered(predict_model_df$CO2_C, levels = c("low", "med", "high"))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis', header=FALSE, quiet = TRUE, include = FALSE}
mixclmod_predict <- predict(mixcl2, predict_model_df, type = "class", Hess = TRUE)
mixclmod_predict
mixclmodConfusion <- confusionMatrix(mixclmod_predict, as.factor(predict_model_df))
```


